### 2019-07-10 paper.tex
Add files via upload

% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
%
% acm templates v1.57 (sigconf template)
%
%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
\settopmatter{printacmref=true}
% mandatory for CCS'19
\usepackage{balance}
% for creating a balanced last page (usually last page with references)

% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Rights management information. 
% This information is sent to you when you complete the rights form.
% These commands have SAMPLE values in them; it is your responsibility as an author to replace
% the commands and values with those provided to you when you complete the rights form.
%
% These commands are for a PROCEEDINGS abstract or paper.

\copyrightyear{2019} 
\acmYear{2019} 
\setcopyright{acmcopyright}
\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
\acmPrice{15.00}
\acmDOI{10.1145/3319535.3339818}
\acmISBN{978-1-4503-6747-9/19/11}

%
% These commands are for a JOURNAL article.
%\setcopyright{acmcopyright}
%\acmJournal{TOG}
%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
%\acmDOI{10.1145/1122445.1122456}

%
% Submission ID. 
% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
% of the event, and this ID should be used as the parameter to this command.
%\acmSubmissionID{123-A56-BU3}

%
% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
% the next command will enable that style.
%\citestyle{acmauthoryear}

 \makeatletter
 % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
 %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
 %                       \abovedisplayshortskip 0pt \@plus2\p@%
 %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
 %                       \@startsection{section}{1}{\z@}%
 %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
 %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
 %                        {\normalfont\large\bfseries\boldmath
 %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
 \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
                        {\normalfont\Large\bfseries\boldmath
                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
                        {\normalfont\normalsize\bfseries\boldmath}}
% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
%                       {\normalfont\normalsize\bfseries}}
 \makeatother

% *** GRAPHICS RELATED PACKAGES ***
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{listings}
% declare the path(s) where your graphic files are
\graphicspath{{../figures/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}

%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
\usepackage{minted}
% *** MATH PACKAGES ***
\usepackage{amsmath}

% *** SPECIALIZED LIST PACKAGES ***
\usepackage{algorithmic}

% *** ALIGNMENT PACKAGES ***
\usepackage{array}

\usepackage{url}

\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
\addbibresource{../library.bib} 
\addbibresource{../cryptobib/abbrev2.bib}
\addbibresource{../cryptobib/crypto_crossref.bib}

% Used for Theorems and Definitions
\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{remark}{Remark}

\usepackage{xcolor} 
\usepackage{xspace}
\usepackage{xstring} % required by IfEqCase

\usepackage{booktabs} % better tables

% ishaq: adding our macros
\input{macros.tex}

\sloppy
\begin{document}
\fancyhead{}
% do not delete this code.

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title[Scheduling and Parallelization for MPC]{Scheduling and Parallelization for MPC}

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author{Muhammad Ishaq}
\authornote{This work was done in part while the author was at RPI.}
\email{m.ishaq@ed.ac.uk}
%\orcid{1234-5678-9012}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
    \institution{University of Edinburgh}
    %\streetaddress{P.O. Box 1212}
    \city{Edinburgh}
    \state{Scotland}
%    \postcode{43017-6221}
}

\author{Lindsey Kennard}
\email{kennal@rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

\author{Ana L. Milanova}
\email{milanova@cs.rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

\author{Vassilis Zikas}
\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
\email{vzikas@inf.ed.ac.uk}
%\orcid{1234-5678-9012}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
    \institution{University of Edinburgh} 
    %\streetaddress{P.O. Box 1212}
    \city{Edinburgh}
    \state{Scotland}
    %    \postcode{43017-6221}
}

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}

\end{abstract}

%
% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10010124.10010138.10010143</concept_id>
<concept_desc>Theory of computation~Program analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003777.10003789</concept_id>
<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10002979</concept_id>
<concept_desc>Security and privacy~Cryptography</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Program analysis}
\ccsdesc[500]{Theory of computation~Cryptographic protocols}
\ccsdesc[300]{Security and privacy~Cryptography}

%
% Keywords. The author(s) should pick words that accurately describe the work being
% presented. Separate the keywords with commas.
\keywords{protocol mixing; linear programming; multiparty computation; program analysis; cryptography}

\begin{comment}
%
% A "teaser" image appears between the author and affiliation information and the body 
% of the document, and typically spans the page. 
\begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{Seattle Mariners at Spring Training, 2010.}
    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
    \label{fig:teaser}
\end{teaserfigure}
\end{comment}
%
% This command processes the author and affiliation and title information and builds
% the first part of the formatted document.
\maketitle

\section{Introduction}
\label{sec:introduction}

In this short paper we define the problem of \emph{optimal parallelization} for Multi-party Computation (MPC), 
set it apart from classical work on parallelization for High-performance Computing (HPC), present preliminary 
results, and set directions for future research.

\ana{1. Intro to MPC (from a PL point of view). Define the problem.}

\ana{2. Difference with HPC. Why "optimal HPC" = "minimal delay" is NOT "optimal MPC"}

\ana{3. Results statement.}

\ana{4. Future directions.}

% no \IEEEPARstart

\section{Scheduling in MPC}

For this treatment we make the following simplifying assumptions:

\begin{enumerate}
\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion.
\item All MPC instructions have the same unit cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.
\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption, 
2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled 
in parallel still cost 2.
\end{enumerate}

\paragraph{Problem 1:} Given a dependence graph of MPC instructions, find the \emph{loop-body} schedule 
with \emph{minimal cost}.

%\begin{figure*}[tbhp]
\begin{figure*}[tbhp]
\begin{tabular}{llll}
\begin{minipage}[b]{4.25cm}

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
for (int i=1; i<=N; i++) {
  A: A[i] = f_1(B[i]);
  B: B[i] = f_2(A[i],D[i-1]);
  C: C[i] = f_3(A[i],D[i-1]);
  D: D[i] = f_4(B[i],C[i]);
}
\end{minted}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.5\textwidth]{figs/dependenceGraph.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[height=0.15\textheight]{figs/MPCloop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[height=0.20\textheight]{figs/MPCaccrossLoop.pdf}
\end{minipage}
\\

(a) Source 
&
(b) Dependence graph
&
(c) Loop-body schedule
&
(d) Across-loop schedule

\end{tabular}
\caption{(a) shows the source code, taken from Aiken and Nicolau~\cite{Aiken:1988}, 
(b) shows the dependence graph. The solid edges denote within-iteration ordering constraints, e.g., A must run before B
because B depends on A. The dashed edges denote across-iteration ordering constraints, e.g., B(2), i.e., B in iteration 2
depends on D(1), i.e., D in iteration 1. (c) shows the optimal loop-body schedule, and (d) shows the
optimal across-loop schedule. A(1), B(1), etc. denote instruction A in iteration 1, etc.}
\label{fig:MPCexample}\vspace{-2ex}
\end{figure*}

\paragraph{Problem 2:} Given a dependence graph of MPC instructions representing a loop body (or given a loop-body schedule), 
find the \emph{across-loop} schedule with  \emph{minimal cost}.

\figref{fig:MPCexample} illustrates the problem with a code example from the classical high-performance computing (HPC) literature. 
\figref{fig:MPCexample}(c) shows the optimal loop-body schedule (answer to Problem 1). Instructions B and C can be scheduled in parallel, 
however, the schedule does not necessarily benefit form parallelization/amortization. If B and C (functions $f_1$ and $f_2$) are the \emph{same}
MPC instruction (e.g., both are MUX, MUL, ADD, or CMP, etc.), then they \emph{can} benefit from parallelization/amortization. Given
our assumptions stated above, the cost of running the loop-body schedule is 3. However, if B and C are different MPC-instructions,
e.g., B is MUL, but C is MUX, then the cost of running the schedule is 4.

\figref{fig:MPCexample}(d) shows the optimal across-loop schedule (answer to Problem 2). We can schedule all A-instructions, across all iterations, 
in parallel, as A depends only on the initial value of B. We cannot do better but schedule B, C, and D sequentially, as they each depend on values
computed in the previous iteration. Again, if B and C are the same instruction, the optimal across-loop schedule will have cost of $1 + 2N$, 
and if B and C are NOT the same instruction, then it will have cost $1+3N$.

\section{Scheduling in HPC}

In this section, we state the classical HPC scheduling problem, and argue the MPC scheduling problem poses unique challenges.

Similar problems arise in the HPC setting.

%\begin{figure*}[tbhp]
\begin{figure*}[tbhp]
\begin{tabular}{lll}

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.4\textwidth]{figs/HPCloop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.4\textwidth]{figs/HPCdoacrossLoop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.6\textwidth]{figs/HPCgreedyLoop.pdf}
\end{minipage}

\\

(a) Loop-body schedule
&
(b) Doacross across-loop schedule
& 
(c) Greedy across-loop schedule

\end{tabular}
\caption{Optimal HPC schedules.}
\label{fig:HPCexample}\vspace{-2ex}
\end{figure*}

\paragraph{Problem 1:} Given a loop body (i.e., dependence graph), find the loop-body schedule with \emph{minimal delay}.

\paragraph{Problem 2:} Given a loop body (or a loop-body schedule), find the \emph{doacross} schedule with minimal delay.

\paragraph{Problem 3:} Given a loop body (or a loop-body schedule), find the \emph{greedy schedule} with minimal delay.

\figref{fig:HPCexample} illustrates the problem in HPC. \figref{fig:HPCexample}(a) shows the optimal loop-body schedule
which results in cost of 3. Again, we assume that all instructions have the same cost, a unit of 1. The key difference with the 
MPC schedule is that regardless of whether B and C are the same instruction opcode or not, B and C can be scheduled 
to run in parallel and the overall running time benefits from parallelization.

The HPC literature has considered different kinds of across-loop scheduling. One scheduling, called \emph{doacross}~\cite{aiken,cytron},
requires that all instructions in an iteration are scheduled on the same processor. The optimal doacross schedule for our example
is shown in \figref{fig:HPCexample}(b) (taken directly from~\cite{aiken}). Aiken and Nicolau proposed a new schedule,
which they called \emph{greedy} schedule, where instructions in an iteration can be scheduled on different processors. 
The optimal greedy schedule for our example, which achieves better minimal delay than the optimal doaccross schedule, 
is shown in \figref{fig:HPCexample}(c) (again, taken directly from~\cite{aiken}).

\ana{Here we have to make the argument that MPC is different from HPC.}
The example illustrates that optimal MPC scheduling and optimal HPC scheduling are different problems.

%\section{\bf Preliminaries}
%\label{sec:preliminaries}

\ana{Relevant papers/books:

\begin{enumerate}

\item Book: "Scheduling and automatic parallelization", Springer 2000, by Alain Darte, Yves Robert and Frederic Vivien. 
This book describes scheduling and parallelization in HPC: given a dependence graph and unlimited 
number of processors, one can compute an optimal (minimal delay) schedule in polynomial time. 
Given a finite number of processors, computing an optimal schedule becomes NP-hard.

\item https://theory.stanford.edu/~aiken/publications/papers/pldi88.pdf. "Optimal loop parallelization", PLDI 1988, by
Alex Aiken and Alex Nicolau. 
Given a dependence graph of the loop body, computes an optimal parallelization across loop iterations. 
The optimality metric is minimal delay. E.g., assuming 2 iterations, the delay is measured by the number 
of instructions in iteration 2 that are lagging behind iteration 1.

\item https://www.cs.indiana.edu/~achauhan/Teaching/B629/2006-Fall/CourseMaterial/1998-popl-knobe-arrayssa.pdf.
"Array SSA", POPL 1998, by Kathleen Knobe and Vivek Sarkar. This paper defines an SSA form for arrays which I think 
is crucial for parallelization across loop iterations.

\end{enumerate}
}

\section{Loop Body Parallelization}
\label{sec:parallelization_within_loop}
%\input{../sections/analysis.tex}

We first consider the answer to Problem 1 (in MPC). 
We argue that optimal parallelization within a loop body is NP-complete (under the above stated assumptions). 
Therefore, compilers must resort to heuristics to compute a schedule for the instructions within a loop body.

We consider two operations, call them $A$ and $M$. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

We make the following assumptions:

\begin{enumerate}

\item $A$ and $M$ are of equal cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single $A$-instrution (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
(or $M$-instructions), namely 1 unit.

\end{enumerate}

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
More precisely, the loop body is such that its instructions can be grouped into such sequences. 
$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
meaning that the source node must execute before the target node): 
\begin{enumerate}
\item $A \rightarrow M \rightarrow A$
\item $A \rightarrow A \rightarrow A$
\item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.

The cost of a schedule $P$ is 
\[\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)\]
where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
a schedule with minimal cost for the sequences above is
\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
The cost of this schedule is 5.

The problem of finding a schedule with minimal cost 
easily reduces to the problem of finding a \emph{shortest common supersequence}, 
a known NP-complete problem \ana{citation needed}. To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)

\ana{Question: are these two assumptions too strong?}

\ana{Assumption 1 can be relaxed --- e.g., if $A$ is 5 times cheaper than $M$, 
we can construct a program where all $A$'s are sequences of five consecutive 
$A$'s, which reduces to the previous problem.}

\ana{Assumption 2, I am not sure if this is acceptable, or maybe it is too strong. 
My intuition is that incorporating into the cost a dependence on the number 
of instructions in each parallel node, is going to make the problem harder, not easier...}

\section{Optimal Loop Parallelization} 
\label{sec:optimal_loop_parallelization}
%\input{../sections/optimal_loop_parallelization.tex}

\ana{Aiken's and Array SSA to define the notion of optimal loop parallelization for MPC.}

We now consider the answer to Problem 2 (in MPC).

\subsection{Problem Overview}

Recall Problem 2, given a loop-body schedule, compute the across-loop schedule
with minimal cost. The following examples illustrate. The first example is the inner 
product computation, and the second is Aiken and Nicolau's example.

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
A = ...
B = ...
ip = 0;
for (i=1; i<=N; i++) {
  t=A[i]*B[i];
  ip=ip+t;
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
A = ...
B = ...
ip = 0;
for (i=1; i<=N; i++) {
  if (i%step5==1) para(t[i],...,t[i+step])
  if (i%step6==1) ip = ip + t[i];
}  
\end{lstlisting}

\end{tabular}

\ana{More to introduce what the schedule is.}

In the above example {\sf step5} is N, meaning that all N multiplications are executed in parallel.
In contrast, {\sf step6} is 1, meaning that the summation statements are executed sequentially.

A more interesting example involves array reads and writes. We return to the Aiken and Nicolau
example:

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  A[i]=f1(B[i]);
  B[i]=f2(A[i],D[i-1]);
  C[i]=f3(A[i],D[i-1]);
  D[i]=f4(B[i],C[i]);
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  if (i%N==1) para(A[i]=f1(B[i]),...);
  if (i%1==1) para(B[i]=f2(A[i],D[i-1]),...);
  if (i%1==1) para(C[i]=f3(A[i],D[i-1]),...);
  if (i%1==1) para(D[i]=f4(B[i],C[i]),...);
}	
\end{lstlisting}

\end{tabular}

The above across-loop schedule means that the first instruction is parallelized---all
N instructions A can be computed at once because each depends only on {\sf B[i]}, which
is assigned outside of the loop. In contrast, subsequent instructions are executed 
sequentially. Take B, for example. There is a dependence ``cycle'' from {\sf B[i]} 
to {\sf D[i]} then {\sf B[i+1]} in the subsequent iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) 
{\sf B[i]}'s cannot be computed in parallel because each one depends on the {\sf B[i]} 
value computed in the previous iteration.

Now consider the above code but slightly modified. {\sf B[i]} and {\sf C[i]} are now 
computed based on the value of {\sf D[i]} from two iterations ago.

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  A[i]=f1(B[i]);
  B[i]=f2(A[i],D[i-2]);
  C[i]=f3(A[i],D[i-2]);
  D[i]=f4(B[i],C[i]);
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  if (i%N==1) para(A[i]=f1(B[i]));
  if (i%2==1) para(B[i]=f2(A[i],D[i-2]));
  if (i%2==1) para(C[i]=f3(A[i],D[i-2]));
  if (i%2==1) para(D[i]=f4(B[i],C[i]));
}	
\end{lstlisting}

\end{tabular}

{\sf B[i]} still depends on values computed in earlier iterations, but one of the values, D[i-2] 
is two iterations earlier. Thus, we can take parallelization steps of 2, e.g., computing 
{\sf B[1]} and {\sf B[2]} in parallel, then {\sf B[3]} and {\sf B[4]} in parallel, etc.
Note however, that we cannot take a parallelization step larger than 2.
The value of B will depend on a value being computed in parallel.

To simplify, we assume that a loop-body schedule is a straight line. \ana{What does this mean?}
\ana{Will we relax?}
The problem therefore becomes: compute the maximal parallelization 
step allowed for each statement. This will account for an MPC schedule with
minimal cost. \ana{Why?}

\subsection{Preliminaries}

\ana{1. MPC-Source}

%\begin{figure*}[tbhp]
\begin{figure}[tbhp]
\begin{tabular}{ll}
\begin{minipage}[b]{4.25cm}

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
// returns val%mod    
int rem(int val, int mod) {
  int rem = 0;
  for (int j = LEN-1; j |$\ge$| 0; j--) 
  {
    rem = rem << 1;
    // rem[0] = val[j]                                                                                                            
    rem = rem + ((val>>j)&1);   
    if (rem |$\ge$| mod) 
    {
      rem = rem - mod;
    }
  }    
  return rem;
}
\end{minted}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
// begin inlined rem
int rem0 = 0;
for (int j = LEN-1; j >= 0; j--) 
{
  rem1 = (j==LEN-1) ? rem0 : rem5;
  rem2 = rem1 << 1;
  rem3 = rem2 + (x1>>j)&1; 
  rem4 = rem3 - y1;
  cnd1 = CMP(rem3 >= y1);
  rem5 = MUX(rem3,rem4,cnd1);
}  
// end inline rem        
\end{minted}
\end{minipage}
\\

(a) Source 
&
(c) MPC-source

\end{tabular}
\caption{(a) shows standard (IMP-like) source for an MPC-friendly implementation 
of integer division~\cite{Demmler}, and (b) shows MPC-source. Line 5 {\sf rem1 = (j==LEN-1) ? rem0 : rem5;} 
is what we call a {\em pseudo-$\phi$} node. Note that the {\sf if}-statement on left
becomes straight-line code in MPC. The program \emph{runs both branches}, and 
at the end runs the conditional of the if-statement followed by the multiplexer. 
If {\sf cnd1} is {\sf true}, then {\sf rem} gets assigned the value along the left branch;
otherwise, it gets assigned the value along the right branch.}
    \label{fig:example}\vspace{-2ex}
\end{figure}

\paragraph{MPC-Source.} MPC-Source is an SSA-based intermediate representation
of MPC programs~\cite{ccs paper}. MPC-Source is, essentially, a straight-line sequence
of ``components'', where each ``component'' is either a nested for-loop sequence (called for-loop block), 
or a simple MPC-statement. \figref{fig:example} shows an example. MPC-Source can be constructed
in a standard way from an SSA-translation of the source---each branch of an 
{\sf if}-statement is added to the sequence, the conditional becomes an MPC comparison (CMP), 
and the $\phi$-node becomes an MPC multiplexer (MUX).

\paragraph{Array SSA.} The other essential concept we build upon is Array SSA. Originally proposed

\ana{3. MPC-Source with Array SSA}

\begin{lstlisting}[language=Java,escapeinside=`']
for (i=1; i<=N; i++) {

  Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise

  A1 = `$\phi$'(A3,A0); @A1 = max(A3,A0);  
  B1 = `$\phi$'(B3,B0); @B1 = max(B3,B0);
  C1 = `$\phi$'(C3,C0); @C1 = max(C3,C0);
  D1 = `$\phi$'(D3,D0); @D1 = max(D3,D0);

  A2[i] = f1(B1[i]); @A2[i] = i;
  A3 = `$\phi$'(A2,A1); @A3 = max(A2, A1);
  B2[i] = f2(A3[i],D1[i-1]); @B2[i] = i;
  B3 = `$\phi$'(B2,B1); @B3 = max(B2, B1);
  C3[i] = f3(A3[i],D1[i-1]); @C2[i] = i;
  C3 = `$\phi$'(C2,C1); @C3 = max(C2, C1);
  D3[i] = f4(B3[i],C3[i]); @D2[i] = i;
  D3 = `$\phi$'(D2,D1); @D3 = max(D2, D1); 
}
\end{lstlisting}

"Index invariant" makes an assertion about when A[i] is written as a function of i.
\ana{Figure out how to reason about cycles based on Array SSA.}

\subsection{Algorithm}

for a statement x = f(y,z), follow the reverse def-use link.

\section{Future Work}
\label{sec:implementation_and_benchmarks}
%\input{../sections/implementation_and_benchmarks.tex}
%\input{../sections/evaluation.tex}

\section{Conclusions}
\label{sec:conclusion}
%\input{../sections/conclusion.tex}

%\begin{acks}
%    \ishaq{TODO}
%\end{acks}

\printbibliography

% that's all folks
\end{document}