### 2022-04-28 overleaf/related.tex

\section{Related Work}
\label{sec:related}

\paragraph{MPC languages and compilers.}

Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.

While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and
AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits.
Our focus is at the intermediate level.

%{\begin{center}
%\includegraphics[width=0.5\linewidth]{figs/focus.png}
%\end{center}
%}

Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of
compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement
different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the
computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution
of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing,
e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.

Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
restrict flow between secure and insecure parts of the program.
%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we
%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our
%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations
%thrust of the proposal.

%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.

\paragraph{Classical HPC compilers.}
Automatic vectorization is a longstanding problem in high-performance computing (HPC).
There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
MPC Source vectorization.
We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
result of the condition at runtime.
The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
can lead to \emph{slowdown}.
%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}

We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
results in HPC.
Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
Vectorization meshes in with those additional optimizations in non-trivial ways.

Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization
at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.