### 2022-05-01 overleaf/results.tex
evaluation section nuking references to biometric_fast

\section{Experimental Results}
\label{sec:results}

\subsection{Experiment Setup}\label{sec:experiment_setup}
We tested our framework with several benchmarks. For the multiparty computation (MPC), we restriced our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources. We stress that there is no such inherent restriction in our framework. We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Netowrk (LAN) and Wide Area Network (WAN). In the LAN setting, we use {\tt c6525-25g} machines for both parties. These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM. The connection between these machines had 10Gbps bandwidth and sub-millisecond latency. This setting reflects typical LAN usecase considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks. In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second. The {\tt c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM. We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms. At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN usecase.

We run all experiments 5 times and report average values of various metrics. Note that standard deviation --- shown as errorbar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.

\subsection{Benchmarks}\label{sec:benchmarks_description}
In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. We used the following benchmarks in our evaluation:

\ishaq{We don't have space to show individual benchmark graphs, and vec data is only plotted in individual benchmark graphs, so specifying what we use for vec experiment in the following text is unnecessary. (unless, we change the graph)}

\ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}

\ishaq{TODO: add a large table for LAN, incorporating as much data as possible}

\begin{enumerate}
    \item {\em Biometric Matching} Server has a database S of {\tt N} records, each record's dimension is {\tt D}. Client submits a query C, client and server compute the closest record to C in an MPC. We use {\tt N=128} for {\em both} and {\tt N=4096} for {\em vec}. {\tt D} is fixed at 4.

    % \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \cite{Demmler:2015}. Parameters are same as above.

    \item {\em Convex Hull} given a polygon of {\tt N} vertices (split between Alice and Bob), convex hull is computed in an MPC. adapted from \cite{Farzan:2021}. We use {\tt N=32} for {\em both} experiment and {\tt N=256} for {\em vec} experiment.

    \item {\em Count 102} Alice has a string of length {\tt N} of symbols, Bob has a regular expression of the form 1(0*)2, together they compute number of substrings that match the regular expression. adapted from \cite{Farzan:2021}. We use {\tt N=1024} for {\em both} and {\tt N=4096} for {\em vec}.

    \item {\em Count 10} Same as {\em Count 102} except now the regular expression is of the form 1(0+). Parameters are same as above.

    \item {\em Cryptonets Max Pooling} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\cite{Dowlin:2016}. We use {\tt rows=64, cols=64} for {\em both} experiment.

    \item {\em Database Join} given two databases with {\tt A} and {\tt B} 2-element records, compute cross join. We use {\tt A=B=32} for {\em both} and {\tt A=B=64} for {\em vec}.

    \item {\em Database Variance} given a database of {\tt len} records, compute variance. \ishaq{Ben/Ana: Please correct the description, I don't know what this benchmark is} We use {\tt len=512} for {\em both} and {\tt len=4096} for {\em vec}.

    \item {\em Histogram} given {\tt N} 5-star ratings, compute their histogram, taken from \cite{Ishaq:2019, Farzan:2021}. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.

    \item {\em Inner Product} given two vectors, each of {\tt N} elements, compute their inner product. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.

    \item {\em k-means Iteration} performs the iteration subroutine of k-means database clustering operation \cite{Jagannathan:2005, Vaidya:2003}. Here {\tt len1} is the size of input data, and {\tt len2} is the number of clusters. We use {\tt len1=32, len2=5} for {\em both} and {\tt len1=256, len2=8} for {\em vec}. \ishaq{Ana/Ben: please update description}

    \item {\em Longest 102} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression. We use same parameters as {\em Count 102}, adapted from \cite{Farzan:2021}.

    \item {\em Max Distance b/w Symbols} Alice has a string of {\tt N} symbols and Bob has some symbol {\tt 0}. The MPC computes the maximum distance between {\tt 0}s in the string. We adapted it from \cite{Farzan:2021}. We use {\tt N=1024} for {\em both} and {\tt N=2048} for {\em vec}.

    \item {\em Minimal Points} Given a set of {\tt N} points (split between Alice and Bob), a set of minimal points is computed i.e. there is no other point that has both a lower x and y coordinate, adapted from \cite{Farzan:2021}. We use {\tt N=32} for {\em both} and {\tt N=64} for {\em vec}.

    \item {\em MNIST ReLU} given an input of ${\tt outer}\times{\tt inner}$ elements, executes the MNIST ReLU subroutine. We use {\tt inner=512} for {\em both} and {\tt inner=2048} for {\em vec}. {\tt outer} is fixed at 16. \ishaq{Ben/Ana, please update}

    \item {\em Private Set Intersection (PSI)} Alice holds set S1 with size {\tt SA}, Bob holds set S2 with size {\tt SB}, together they compute intersection of their sets. We use {\tt SA=SB=128} for {\em both} and {\tt SA=SB=1024} for {\em vec}.
\end{enumerate}

\subsection{Results and Analysis}\label{sec:result_analysis}

   \begin{table*}[htbp]
       \centering
       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
       \label{table:metrics}
       \rowcolors{5}{}{gray!10}
       \tabcolsep3pt
       \begin{tabular}{lrrrrrrrrrrrr}
           \toprule
           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
          \midrule
            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
            \midrule
            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
            \midrule
            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
            \midrule
            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
            \midrule
            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
            \midrule
            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
            \midrule
            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
            \midrule
            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
            \midrule
            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
            \midrule
            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
            \midrule
            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
            \midrule
            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
            \midrule
            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
            \midrule
            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
            \midrule
            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
          \bottomrule
       \end{tabular}
       \label{table:metrics}
   \end{table*}

\begin{figure*}[htbp]
\centering
%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\input{graphs/lan/all-hist-OnlineSetupTimesec}
\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
\label{fig:graph_all_eval_time}
\end{figure*}

A detailed summery of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection. These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of inner product and biometric matching. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols). There is very little parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, longest 102 and max distance between symbols suffer a slowdown of 0.5x. The structure of these benchmarks is such that transformation to vectorized code increases multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. We fix this via a simple heuristic where if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we reject the transformation. Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time. Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments. As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

We present evaluation for communication size in \cref{fig:graph_comm_size_time}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively. \ishaq{Ana: Let me know if I should add more substance.}

\begin{figure}[htbp]
\centering
\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
\label{fig:graph_biometric_eval_time}
\end{figure}

\begin{figure}[htbp]
\centering
\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
\caption{Biometric Matching Communication Size, x-axis lists database size}
\label{fig:graph_biometic_comm_size}
\end{figure}

\begin{figure}[htbp]
\centering
\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
\caption{Biometric Matching Circuit Generation Time, x-axis lists database size}
\label{fig:graph_biometic_circ_gen_time}
\end{figure}

Next, we look closely at circuit evaluation \cref{fig:graph_biometric_eval_time}, communication size \cref{fig:graph_biometic_comm_size} and circuit generation time \cref{fig:graph_biometic_circ_gen_time} for biometric matching benchmark. For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation. Consequently, non-vectorized bars are missing beyond this threshold. Notice that vectorization improves all metrics. A database of size {\tt N=128} is smaller than any real world usecase, therefore lets use it as baseline for improvement. Comparing performance improvement between BMR and GMW, we see more speedup for BMR (23x vs 10x), GMW gets more communication size reduction (10x vs 2.5x) and circuit generation sees a speedup of 35x and 45x for BMR and GMW respectively.

\begin{figure}[htbp]
\centering
\resizebox{3.6in}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
\caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
\label{fig:graph_comparison_eval_time}
\end{figure}

Since our vectorization framework is network agnostic, it produces the same circuit for both LAN and WAN. This means that the number of gates and communication size remain the same. Moreover, time for circuit generation, which is a local operation, also remains unchanged. Setup and Online times, however, increase due to lower bandwidth and higher latency of the WAN. Indeed, this is what we observe in \cref{fig:graph_comparison_eval_time}.

\begin{figure*}[htbp]
\centering
\input{graphs/lan/all-hist-CommunicationMiB}
\caption{Communication Size of Benchmarks}
\label{fig:graph_comm_size}
\end{figure*}

\begin{figure*}[htbp]
\centering
\input{graphs/lan/all-hist-CircuitGenerationTimesec}
\caption{Circuit Generation Time of Benchmarks}
\label{fig:graph_circ_gen_time}
\end{figure*}

\begin{figure*}[htbp]
\centering
\input{graphs/lan/all-hist-TotalGates}
\caption{Number of Gates of Benchmarks}
\label{fig:graph_total_gates}
\end{figure*}

\begin{figure*}[htbp]
\centering
\input{graphs/lan/all-hist-OnlineTimesec}
\caption{Online Time of Benchmarks}
\label{fig:graph_online_time}
\end{figure*}

\begin{figure*}[htbp]
\centering
\input{graphs/lan/all-hist-SetupTimesec}
\caption{Setup Time of Benchmarks}
\label{fig:graph_setup_time}
\end{figure*}

\subsection{Discussion}
\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}

\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}